\bf Landslide: \\ Systematic Dynamic Race Detection in Kernel Space

### 2012-05-13 evaluation.tex
more

\revision{In evaluating Landslide, we sought to answer seven overall questions.

@@

Many problems can arise from this error. The one Landslide discovered was as follows: During this gap, the child can take its own parent pointer mutex (after the pointer was changed to init), and attempt to move itself from init's ``live child-list'' to the ``dead child-list'' before the parent moved the child to init's child-list at all. This results in the child instead moving the {\em shell process} off of init's ``live child-list''. Init gets signalled, reaps the reparented child, does \texttt{wait} again (expecting the shell to be alive), which fails. This results in init infinitely looping around failing calls to \texttt{wait}. To catch this bug, we added an assert statement in the \texttt{wait} failure case that the current thread was not init (because init should always have children to reap or wait for), which Landslide then caused to trip.

@@

\begin{itemize}
	\item All Landslide trial times include the Simics start-up and kernel boot-up time (time between issuing the command and the test case beginning to run), roughly 15 seconds for POBBLES and 10 seconds for LudicrOS.
	\item All Landslide trials were run on the Gates-Hillman cluster machines (2.6 GHz Xeon).
	\item All Landslide trials were run with ``backwards exploration'' enabled (Section~\ref{sec:using-search}).
	\item All trials were also run with Landslide configured to pay attention to only the relevant system calls (using \texttt{within\_function}; Section~\ref{sec:using-decision}).

@@

\begin{enumerate}
	\item {\bf Fewer is faster.} While it is theoretically possible that a tree built of finer-grained interleavings might encounter a bug after fewer overall backtracks, we found that this did not happen in practice.\footnote{
		The one suspicious case is the LudicrOS vanish/vanish bug - exploring the ``both'' tree found the bug faster than exploring the ``unlock'' tree, despite the latter's decision points being a subset of the former's. However, we see that exploring the ``lock'' tree found the bug faster than either, so overall the ``both'' tree did not outperform the fastest of the smaller trees.}
		In general, for two sets of decision points that both find the same bug, the one that results in shorter {\em branches} will result in fewer {\em backtracks} needed to expose the bug, and hence less overall time.
	\item {\bf Different decision points are differently likely to expose different bugs.} We see that even though the ``lock'' and ``unlock'' trees tended to be about the same size, they were each sometimes better than the other at finding bugs. The ``lock'' tree did better on the vanish/vanish bugs and the yield/vanish bug, while the ``unlock'' tree did better on the wait/wait bug and the thread\_fork/vanish bug.\footnote{
	The latter two bugs needed no more than the default set to uncover at all, but we claim this still demonstrates the principle in general.}
	\item {\bf Finding a bug is fast, if it exists.} As especially exhibited in the POBBLES vanish/vanish (b) case, if two sets of decision points generate trees of approximately equal size, but one tree contains a bug and the other doesn't, then searching the bugful tree will likely terminate much more quickly.
	Of course, it is always possible that a bug may only exist in the very last branch of a tree, but we found that in general bugs show up early during exploration.
\end{enumerate}

>>>>

\revision{In evaluating Landslide, we sought to answer two overall questions.

@@

Many problems can arise from this error. The one Landslide discovered was as follows: During this gap, the child can take its own parent pointer mutex (after the pointer was changed to init), and attempt to move itself from init's ``live child-list'' to the ``dead child-list'' before the parent moved the child to init's child-list at all. This results in the child instead moving the {\em shell process} off of init's ``live child-list''. Init gets signalled, reaps the reparented child, does \texttt{wait} again (expecting the shell to be alive), which fails. This results in init infinitely looping around failing calls to \texttt{wait}. To catch this bug \revision{immediately, rather than relying on infinite loop detection,} we added an assert statement in the \texttt{wait} failure case that the current thread was not init (because init should always have children to reap or wait for), which Landslide then caused to trip.

@@

\begin{itemize}
	\item All Landslide trial times include the Simics start-up and kernel boot-up time (time between issuing the command and the test case beginning to run), roughly 15 seconds for POBBLES and 10 seconds for LudicrOS.
	\item All Landslide trials were run on the Gates-Hillman cluster machines (2.6 GHz Xeon; \revision{four cores, though only one was used; 8 GB RAM}).
	\item All Landslide trials were run with ``backwards exploration'' enabled (Section~\ref{sec:using-search}).
	\item All trials were also run with Landslide configured to pay attention to only the relevant system calls (using \texttt{within\_function}; Section~\ref{sec:using-decision}).

@@

\begin{enumerate}
	\item {\bf Fewer is faster.} While it is theoretically possible that a tree built of finer-grained interleavings might encounter a bug after fewer overall backtracks, we found that this did not happen in practice.\footnote{
		The one suspicious case is the LudicrOS vanish/vanish bug - exploring the ``both'' tree found the bug faster than exploring the ``unlock'' tree, despite the latter's decision points being a subset of the former's. However, we see that exploring the ``lock'' tree found the bug faster than either, so overall the ``both'' tree did not outperform the fastest of the smaller trees.}
		In general, for two sets of decision points that both find the same bug, the one that results in shorter {\em branches} will result in fewer {\em backtracks} needed to expose the bug, and hence less overall time.
	\item {\bf Different decision points are differently likely to expose different bugs.} We see that even though the ``lock'' and ``unlock'' trees tended to be about the same size, they were each sometimes better than the other at finding bugs. The ``lock'' tree did better on the vanish/vanish bugs and the yield/vanish bug, while the ``unlock'' tree did better on the wait/wait bug and the thread\_fork/vanish bug.\footnote{
	The latter two bugs needed no more than the default set to uncover at all, but we claim this still demonstrates the principle in general.}
	\item {\bf Finding a bug is fast, if it exists.} As especially exhibited in the POBBLES vanish/vanish (b) case, if two sets of decision points generate trees of approximately equal size, but one tree contains a bug and the other doesn't, then searching the bugful tree will likely terminate much more quickly.
	Of course, it is always possible that a bug may exist only in the very last branch of a tree, but we found that in general bugs show up early during exploration.
\end{enumerate}

### 2012-05-13 futurework.tex
more

We believe that a tool such as Landslide could change the way students learn such skills for the better, beyond simply increasing their likelihood of finding bugs during testing. The decision tree is a much more structured way of expressing a test case's concurrent behaviour than the way 15-410 currently teaches, which is ``think hard until you figure out the buggy interleaving''. Having students use Landslide, even to a small extent, might encourage them to think in this more structured way.\footnote{
\revision{In 15-410, before the kernel project, students also implement a user-space thread library. Just as Landslide could help the students during the kernel project, so too could a user-space systematic exploration tool help during the thread library project.}}

@@

Landslide could make use of data race detection techniques\cite{datacollider}, however, to strike a middle ground in which shared memory accesses it chooses.
\begin{enumerate}
	\item It should be aware of the types and interfaces associated with the kernel's synchronisation primitives (mutexes, condvars, semaphores, etc), and be able to treat operations on those as ``guaranteed to work'' (similar to Section~\ref{sec:por-independence}).
	\item Next, it should recognise acquire and release (or sleep and wake) operations on synchronisation primitives, and use that information to track lockset-type information.
	\item Ultimately, it should use the lockset information in conjunction with the happens-before relation to identify which memory accesses are ``raciest'', and add decision points around those.
\end{enumerate}

@@

Major benefit could be achieved by redesigning Landslide to work in a virtualised environment instead of a simulated one. This would pose several challenges, because Landslide would no longer be running in ``god mode'', having immediate access to every single instruction and memory access executed. It would, however, still be able to modify the kernel's memory and address space, which would be important to harness for increased control over the kernel's execution.

@@

Finally, Landslide will need to find new ways to control non-deterministic events in the system. Since virtualised kernels use the same timer interrupt as the host they run on, Landslide is unlikely to be able to intercept it. Instead, we may replace a guest kernel's timer-driven context switcher with a ``Landslide-driven'' context switcher, leaving the timer handler only to do other work.

@@

One member of group 1 from the user study fully explored decision trees from \texttt{double\_wait} and \texttt{vanish\_vanish} (Section~\ref{sec:eval-victory}), and found an oddity in the exploration time for a particular decision set.

>>>>

We believe that a tool such as Landslide could change the way students learn such skills for the better, beyond simply increasing their likelihood of finding bugs during testing. The decision tree is a much more structured way of expressing a test case's concurrent behaviour than the way 15-410 currently teaches, which is ``think hard until you figure out the buggy interleaving.'' Having students use Landslide, even to a small extent, might encourage them to think in this more structured way.\footnote{
\revision{In 15-410, before the kernel project, students also implement a user-space thread library. Just as Landslide could help the students during the kernel project, so too could a user-space systematic exploration tool help during the thread library project.}}

@@

Landslide could make use of data race detection techniques\cite{datacollider}, however, to strike a middle ground in which shared memory accesses it chooses.
\begin{enumerate}
	\item It should be aware of the types and interfaces associated with the kernel's synchronisation primitives (mutexes, condvars, semaphores, etc), and be able to treat operations on those as ``guaranteed to work'' (similar to Section~\ref{sec:por-independence}).\footnote{
	\revision{If such operations were instead themselves incorrect, it would mean that data race detection might fail to identify certain racy accesses. The technique would then still identify some racy accesses, but might miss others.}}
	\item Next, it should recognise acquire and release (or sleep and wake) operations on synchronisation primitives, and use that information to track lockset-type information.
	\item Ultimately, it should use the lockset information in conjunction with the happens-before relation to identify which memory accesses are ``raciest'', and add decision points around those.
\end{enumerate}

@@

Major benefit could be achieved by redesigning Landslide to work in a virtualised environment instead of a simulated one. This would pose several challenges, because Landslide would no longer be running in ``\revision{omniscient} mode'', having immediate access to every single instruction and memory access executed. It would, however, still be able to modify the kernel's memory and address space, which would be important to harness for increased control over the kernel's execution.

@@

Finally, Landslide will need to find new ways to control non-deterministic events in the system. Since virtualised kernels use the same timer interrupt as the host they run on, Landslide \revision{will need to intercept it, and perhaps control the guest kernel's context switching with a custom ``Landslide-driven'' context switcher instead.}

@@

One member of Group 1 from the user study fully explored decision trees from \texttt{double\_wait} and \texttt{vanish\_vanish} (Section~\ref{sec:eval-victory}), and found an oddity in the exploration time for a particular decision set.

### 2012-05-13 testcases.tex
more

The test cases that use the \texttt{thread\_fork} system call are written in assembly, because thread creation can only safely be done from C using a userspace thread library. However, using a userspace thread library for thread creation might forcibly serialise certain operations that would expose kernel bugs only if concurrent, and also necessitates many more system calls, which dramatically increases the size of the decision tree.

>>>>

The test cases that use the \texttt{thread\_fork} system call are written in assembly, because thread creation can only safely be done from C using a user-space thread library. However, using a user-space thread library for thread creation might forcibly serialise certain operations that would expose kernel bugs only if concurrent, and also necessitates many more system calls, which dramatically increases the size of the decision tree.
